---
title: "exploratory_"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Unhash the first line of code before running program for the first time
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)

```

```{r}
rawdata <- read.csv('raw_data.csv')
head(rawdata, 10)
glimpse(rawdata)
nrow(rawdata)
ncol(rawdata)

```
### About the dataset

The dataset comprises of Rows: 
- 72,531, Columns: 147 
- 16 dummy variables (flags) 
- counts of total population, can be grouped by socioeconomic, demographic, car no car etc.

Complete VARIABLE LIST IN NEXT CELL
```{r}
variables_list <- read.csv('variable_list.csv')
variables_list
```
```{r}
# creating first rendition of new dataset, attempting to remove null values
data <- rawdata %>% na.omit %>% distinct()
data <- complete(data)
head(data,100)

data[is.na(data)] <- NaN
glimpse(data)


#there are still null values but these are in demographic shares, will have to deal with them later
```


```{r}
#Number of urban-tracts 

#by state
data %>% group_by(State) %>% summarize(sum(Urban)) 

#by county
data %>% group_by(County) %>% summarize(sum(Urban))

data %>% select(starts_with("LATracts")) %>% select(c(-5)) %>% sum

#looking at the sum of columns of this data we see the results to be
#LATracts_half	LATracts1	LATracts10	LATracts20
#46335	       24362	      3186	        347

```

## Pursuing investigation into LILAs    

LILA = Low income, low access. 
```{r}
# isolating the columns that deal with shares/percentages


datashare <- data %>% select(c("CensusTract", 
                               "Urban", 
                               "MedianFamilyIncome"
"lapophalfshare", 
      
"lalowihalfshare",

'lakidshalfshare',

"laseniorshalfshare",

"lawhitehalfshare",

"lablackhalfshare",

"laasianhalfshare",

"lanhopihalfshare",

"laaianhalfshare",

"laomultirhalfshare",

"lahisphalfshare",

"lahunvhalfshare",

"lasnaphalfshare",

"lapop1share",

"lalowi1share",

"lakids1share",

"laseniors1share",

"lawhite1share",

"lablack1share",

"laasian1share",

"lanhopi1share",

"laaian1share",

"laomultir1share",

"lahisp1share",

"lahunv1share",

"lasnap1share",

"lapop10share",

"lalowi10share",

"lakids10share",

"laseniors10share",

"lawhite10share",

"lablack10share",

"laasian10share",

"lanhopi10share",

"laaian10share",
"laomultir10share",

"lahisp10share",

"lahunv10share",

"lasnap10share",

"lapop20share",

"lalowi20share",

"lakids20share",

"laseniors20share",

"lawhite20share",

"lablack20share",

"laasian20share",

"lanhopi20share",

"laaian20share",

"laomultir20share",

"lahisp20share",

"lahunv20share",

"lasnap20share"))

datashare[ datashare == "NULL"] <- NA
datashare[ complete.cases(datashare),]

head(datashare, 10)

glimpse(datashare)

# This data has Census tract number as the primary identifier to be able to join back to any additional datasets. The share variables are 53 variables from the initial 147
#Now to drop nulls

library("Hmisc")
```

## Looking at some correlation matrices
```{r}
#importing data with no null values, brings the 
corr_data = read.csv('Nonulldata.csv')


glimpse(corr_data)
corr_data[] <- lapply(corr_data, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
sapply(corr_data, class)

corr_data <- corr_data %>% select(c("CensusTract", 
                               
                               "MedianFamilyIncome",
"lapophalfshare", 
      
"lalowihalfshare",

'lakidshalfshare',

"laseniorshalfshare",

"lawhitehalfshare",

"lablackhalfshare",

"laasianhalfshare",

"lanhopihalfshare",

"laaianhalfshare",

"laomultirhalfshare",

"lahisphalfshare",

"lahunvhalfshare",

"lasnaphalfshare",

"lapop1share",

"lalowi1share",

"lakids1share",

"laseniors1share",

"lawhite1share",

"lablack1share",

"laasian1share",

"lanhopi1share",

"laaian1share",

"laomultir1share",

"lahisp1share",

"lahunv1share",

"lasnap1share",

"lapop10share",

"lalowi10share",

"lakids10share",

"laseniors10share",

"lawhite10share",

"lablack10share",

"laasian10share",

"lanhopi10share",

"laaian10share",
"laomultir10share",

"lahisp10share",

"lahunv10share",

"lasnap10share",

"lapop20share",

"lalowi20share",

"lakids20share",

"laseniors20share",

"lawhite20share",

"lablack20share",

"laasian20share",

"lanhopi20share",

"laaian20share",

"laomultir20share",

"lahisp20share",

"lahunv20share",

"lasnap20share"))

M <- cor(corr_data, method = c("pearson"))
glimpse(corr_data)
```


```{r}
library(corrplot)
cplt1 <- corrplot(M, method = 'color', type = 'upper', tl.cex = 0.5)

c <- jpeg("cplt.jpg")
```

